Parameters estimation via HOSVD

### 2024-10-06 Code/R/Research/HOSVD parameters estimate.Rmd
Added HO-ESPRIT algorithms

```{r tensorization}
tens3 <- function(s, I, L) {
  require("rTensor")
  v <- as.vector(s)
  N <- length(v)
  J <- N - I - L + 2
  X <- outer(1:L, 1:J, function(l, j) l + j) |> outer(1:I, function(lj, i) s[lj + i - 2])
  return(as.tensor(X))
}

# diagonal averaging i + j + k = const, const = 3:(I+L+J)
reconstruct.group3 <- function(X.tens) {
  stopifnot(is(X.tens, "Tensor"))
  X <- X.tens@data
  I <- length(X[, 1, 1])
  L <- length(X[1, , 1])
  J <- length(X[1, 1,])
  s <- vector(mode = "numeric", length = I + L + J - 2)
  for (C in 3:(I + L + J)) {
    sum <- 0
    count <- 0
    for (i in 1:(C - 2)) {
      for (l in 1:(C - 1 - i)) {
        if (i <= I && l <= L && C - i - l <= J) {
          sum <- sum + X[i, l, C - i - l]
          count <- count + 1
        }
      }
    }
    s[C - 2] <- sum / count
  }
  return(s)
}
```

@@

>>>>

```{r tensorization}
tens3 <- function(s, I, L, kind = c("HO-SSA", "HO-MSSA")) {
  require("rTensor")
  if (identical(kind[1], "HO-SSA")) {
    N <- length(s)
    J <- N - I - L + 2
    X <- outer(1:I, 1:L, `+`) |>
      outer(1:J, function(il, j)
        s[il + j - 2])
  }
  else if (identical(kind[1], "HO-MSSA")) {
    if (missing(L))
      L <- I
    N <- nrow(s)
    K <- N - L + 1
    Q <- ncol(s)
    X <- apply(s, 2, Rssa::hankel, L = L, simplify = FALSE) |>
      Reduce(cbind, x = _) |>
      rTensor::fold(1, 2:3, modes = c(L, K, Q))
  }
  return(as.tensor(X))
}

reconstruct.group3 <- function(X.tens) {
  stopifnot(is(X.tens, "Tensor"))
  X <- X.tens@data
  I <- length(X[, 1, 1])
  L <- length(X[1, , 1])
  J <- length(X[1, 1,])
  s <- vector(mode = "numeric", length = I + L + J - 2)
  for (C in 3:(I + L + J)) {
    sum <- 0
    count <- 0
    for (i in 1:(C - 2)) {
      for (l in 1:(C - 1 - i)) {
        if (i <= I && l <= L && C - i - l <= J) {
          sum <- sum + X[i, l, C - i - l]
          count <- count + 1
        }
      }
    }
    s[C - 2] <- sum / count
  }
  return(s)
}
```

@@

```{r tensor-based-esprit}
tens_esprit <- function(s, I, L, groups, kind = c("HO-SSA", "HO-MSSA"), est_dim = 3, status = TRUE) {
  if (identical(kind[1], "HO-SSA"))
    H <- tens3(s, I, L)
  else if (identical(kind[1], "HO-MSSA"))
  {
    if (missing(L))
      L <- I
    H <- tens3(s, L, kind = "HO-MSSA")
  }
  max_rank <- max(sapply(groups, max))
  H.hooi <- tucker_mod(H, rep(max_rank, 3), status = status)
  estimates <- list()
  for (i in seq(groups)) {
    U <- H$U[[est_dim]][, groups[[i]], drop = FALSE]
    Z <- solve(U[-nrow(U), ], U[-1, ])
    poles <- eigen(Z, only.values = TRUE)$values
    estimates[[i]] <- list(
      poles = poles,
      damping = -Re(log(poles)),
      freq = Im(log(poles)) / 2 / pi
    )
  }
  estimates
}
```

```{r test-1d-esprit-no-noise}

```

```{r complex-noise-generator}
CCSWGN <- function(n, mean = 0, sd = 1) {
  rnorm(n, mean = Re(mean), sd = sd / sqrt(2)) + 1i * rnorm(n, mean = Im(mean), sd = sd / sqrt(2))
}
```

```{r test-1d-esprit-with-noise}
set.seed(5)
n <- 25
SNR <- 30
signal <- exp((-0.01 + 2i * pi * 0.2) * 0:(n - 1)) + exp((-0.02 + 2i * pi * 0.22) * 0:(n - 1))
sd <- sqrt(mean(abs(signal)^2) / SNR)
noise <- CCSWGN(n, mean = 0, sd = sd)
x <- signal + noise
```
